Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas

### 2022-01-05 analysis/paper/London_HauloutBeahvior.Rmd
REDESIGN & IMPROVE HO WX PLOTS

```{r setup, include = FALSE}
library(targets)
library(conflicted)
library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
library(lubridate)
conflict_prefer("month","lubridate")
library(glue)
library(scales)
library(DBI)
library(dbplyr)
library(RPostgres)
library(pingr)
library(sf)
library(flextable)
library(patchwork)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(ggsci)
library(ggrepel)
library(gghighlight)
library(ggtext)
library(glmmLDTS)
library(mgcv)
library(ragg)

@@

```{r seal-hours-calendar, include=TRUE, fig.id = "dataCal", fig.height = 4.75, fig.cap = "Distribution of hourly percent-dry bio-logger data from 1 March to 15 July for each species. Data are grouped by day-of-year and presented in seal-hours collated across all years between 2005 and 2020. The higher density of data from May and June in ribbon and spotted seals coincides with peak molting when seals (and their attached bio-loggers) are more likely hauled out. Additionally, many bio-logger deployments started in April and May."}

ggplot(data = seal_hours_tbl, aes(x=yday, y=n, fill=n)) +
  geom_col(binwidth = 1) +
  scale_fill_gradient(low = "#6baed6", high = "#08306b") +
  scale_x_continuous(breaks = c(60, 91, 121, 152, 182),
                     labels = c("March", "April", "May", "June", "July"),
                     expand = expansion(mult = c(0.01, 0.025))) +
  guides(fill = guide_legend(title.position = "bottom", title.hjust = 0.5)) +
  facet_grid(species ~ .) + 
  theme(axis.text.x = element_text(hjust = 0),
        axis.text.y = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom") +
  labs(fill = "seal-hours of bio-logger data") +
  ggtitle("Density of haul-out behavior observations varies from March through July",
          subtitle = "data are grouped by day-of-year and presented in seal-hours collated across years (2005-2020)")

@@

@@

Visualizing the marginal effect of each weather covariate on haul-out
probability is difficult in this analysis because of the collinearity
between covariates as well as the temporal variation in covariate values
during the study period. The relationship of each weather covariate with
haul-out probability, averaged over the other weather conditions, is
more variable than simply showing model coefficients would imply. Figure
\@ref(fig:beardedHOwx) presents the predicted haul-out probability of
bearded seals across the range of weather conditions encountered in the
observed data. The collinearity and temporal variation are apparent in
the visual patterns that emerge.

```{r beardedHOwx,fig.height = 5.5, include=TRUE, fig.id = "beardedHOwx", fig.cap="Variability in predicted haul-out probability of bearded seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

p_temp <- ggplot(plot_df, aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df, aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df, aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df, aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

@@

((p_temp + p_wind) / (p_pressure + p_precip)) +
  plot_annotation(
    title = 'Bearded seal haul-out behavior is less influeced by weather')
```

@@

```{r ribbonHOwx, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOwx", fig.cap="Variability in predicted haul-out probability of ribbon seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

p_temp <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  geom_smooth(se = FALSE) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) ),
                 name = "") +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

@@

p_temp + p_wind + p_pressure + p_precip +
  plot_annotation(
    title = 'Ribbon seal haul-out behavior is influenced by weather')
```

@@

```{r spottedHOwx,fig.height = 5.5, include=TRUE, fig.id = "spottedHOwx", fig.cap="Variability in the predicted haul-out probability of spotted seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

library(ggsci)

p_temp <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) ),
                 name = "") +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

@@

p_temp + p_wind + p_pressure + p_precip +
  plot_annotation(
    title = 'Spotted seal haul-out behavior is influenced by weather')
```

>>>>

```{r setup, include = FALSE}
library(targets)
library(conflicted)
library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
library(lubridate)
conflict_prefer("month","lubridate")
library(glue)
library(scales)
library(DBI)
library(dbplyr)
library(RPostgres)
library(pingr)
library(sf)
library(flextable)
library(patchwork)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(ggsci)
library(ggrepel)
library(gghighlight)
library(ggtext)
library(glmmLDTS)
library(mgcv)
library(ragg)
library(stickylabeller)

@@

```{r seal-hours-calendar, include=TRUE, fig.id = "dataCal", fig.height = 4, fig.cap = "Distribution of hourly percent-dry bio-logger data from 1 March to 15 July for each species. Data are grouped by day-of-year and presented in seal-hours collated across all years between 2005 and 2020. The higher density of data from May and June in ribbon and spotted seals coincides with peak molting when seals (and their attached bio-loggers) are more likely hauled out. Additionally, many bio-logger deployments started in April and May."}

# bar chart instead of a strip
#
# ggplot(data = seal_hours_tbl, aes(x=yday, y=n, fill=n)) +
#   geom_col(binwidth = 1) +
#   scale_fill_gradient(low = "#6baed6", high = "#08306b") +
#   scale_x_continuous(breaks = c(60, 91, 121, 152, 182),
#                      labels = c("March", "April", "May", "June", "July"),
#                      expand = expansion(mult = c(0.01, 0.025))) +
#   guides(fill = guide_legend(title.position = "bottom", title.hjust = 0.5)) +
#   facet_grid(species ~ .) + 
#   theme(axis.text.x = element_text(hjust = 0),
#         axis.text.y = element_blank(),
#         axis.title = element_blank(),
#         panel.grid = element_blank(),
#         legend.position = "bottom") +
#   labs(fill = "seal-hours of bio-logger data") +
#   ggtitle("Density of haul-out behavior observations varies from March through July",
#           subtitle = "data are grouped by day-of-year and presented in seal-hours collated across years (2005-2020)")

ggplot(data = seal_hours_tbl, aes(x = yday, y = 1, fill = n))+
        geom_tile()+
        scale_fill_viridis_c(option="mako") +
        scale_x_continuous(breaks = c(60, 91, 121, 152, 182),
                           labels = c("01 March", "01 April", "01 May", "01 June", "01 July"),
                           expand = expansion(mult = c(0, 0))) +
        scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  guides(fill = guide_colorbar(title.position = "bottom", 
                               title.hjust = 0.5, barwidth = 15,
                               barheight = 0.5)) +
  facet_grid(species ~ .) + 
  theme(axis.text.x = element_text(hjust = 0),
        axis.text.y = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom") +
  labs(fill = "seal-hours of bio-logger data") +
  ggtitle("Density of haul-out behavior observations varies from March through July",
          subtitle = "data are grouped by day-of-year and presented in seal-hours collated across years (2005-2020)")

@@

Visualizing the marginal or conditional effect of an individual weather
covariate on haul-out probability is difficult in this analysis because of the
collinearity between covariates as well as the spatial and temporal variation
across such a large region. The relationship of each weather covariate with
haul-out probability, averaged over the other weather conditions, is more
variable than simply showing model coefficients would imply. To elucidate
underlying patterns between the model's predicted haul-out probability and
specific weather covariates, we explored a subset of predictions during three
separate time windows. The same set of predictions used to create the haul-out
probability surfaces was reduced to include only local solar hours between 06:00
and 18:00 during three 7-day time windows. The time windows were centered on the
day of maximum haul-out probability and 25 days before and 25 days after peak
haul out. While this approach does not entirely remove patterns of collinearity
from the visualization (e.g. temperature increases with day of year), by
limiting our range of days and hours we gain some insight into how haul-out
probability varies across different weather covariate values. Figures
\@ref(fig:beardedHOwx), @\ref(fig:ribbonHOwx), and @\ref(fig:spottedHOwx)
present the predicted haul-out probability with respect to age and sex class
across the range of temperature and wind conditions encountered in the three
time windows. The visualizations also include vertical lines representing 95%
confidence intervals around the predicted haul-out probability to better
communicate the variation in model uncertainty. Additionally, simple linear fits
are shown to assist interpretation of overall patterns.

@@

```{r beardedHOwx,fig.height = 5.5, include=TRUE, fig.id = "beardedHOwx", fig.cap="Variability in predicted haul-out probability of bearded seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

plot_df <- bearded_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label = TRUE),
         day = lubridate::day(date)) %>% 
  filter(age_sex != "YOUNG OF YEAR",
         between(solar_hour, 8,16)) %>% 
  group_by(age_sex)

plot_df_early <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg-25)-3, (peak_avg-25)+3))),
         season = "early") %>% 
  unnest(cols = c(data))

plot_df_peak <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg)-3, (peak_avg)+3))),
         season = "peak") %>% 
  unnest(cols = c(data))

plot_df_late <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg+25)-3, (peak_avg+25)+3))),
         season = "late") %>% 
  unnest(cols = c(data))

plot_df <- bind_rows(list(plot_df_early, plot_df_peak, plot_df_late)) %>% 
  mutate(date_start = case_when(
    season == "peak" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 3, "%d %b"),
    season == "early" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 28, "%d %b"),
    season == "late" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 22, "%d %b")
  )) %>% 
  mutate(date_end = case_when(
    season == "peak" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 3, "%d %b"),
    season == "early" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 22, "%d %b"),
    season == "late" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 28, "%d %b")
  )) %>% 
  mutate(season = forcats::fct_relevel(season, "early", "peak", "late"),
         season = forcats::fct_recode(season, "peak haul out - 25 days" = "early",
                             "peak haul out" = "peak",
                             "peak haul out + 25 days" = "late"))

p_temp <- ggplot(plot_df, 
                 aes(temp2*27, ho_prob, color = age_sex, fill = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.1, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, size = 0.3, alpha = 0.1,
              show.legend = FALSE) +
  scale_fill_d3() +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1),
                                      direction = "horizontal")) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  facet_wrap(. ~ season + date_start + date_end, scales = "free_x",
             labeller = label_glue('{season}\n{date_start} - {date_end}')) +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df, aes(wind*10, ho_prob, color = age_sex, fill = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.1, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, size = 0.3, alpha = 0.1,
              show.legend = FALSE) +
  scale_fill_d3() +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 0.75),
                                      direction = "horizontal")) +
  facet_wrap(. ~ season + date_start + date_end, scales = "free_x",
             labeller = label_glue('{season}\n{date_start} - {date_end}')) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  xlab("total wind (m/s)") + ylab("haul-out probability")

@@

p_temp / p_wind +
  plot_annotation(
    title = 'The influence of temperature and wind on predicted bearded seal \nhaul-out behavior varies within season',
    subtitle = 'data are subset to only include local solar hours between 0800 and 1600',
    caption = 'Solid lines represent linear fits to the predicted values and are \nprovided to better indicate marginal patterns.')
```

@@

```{r ribbonHOwx, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOwx", fig.cap="Variability in predicted haul-out probability of ribbon seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

plot_df <- ribbon_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label = TRUE),
         day = lubridate::day(date)) %>% 
  filter(age_sex != "YOUNG OF YEAR",
         between(solar_hour, 8,16)) %>% 
  group_by(age_sex)

plot_df_early <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg-25)-3, (peak_avg-25)+3))),
         season = "early") %>% 
  unnest(cols = c(data))

plot_df_peak <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg)-3, (peak_avg)+3))),
         season = "peak") %>% 
  unnest(cols = c(data))

plot_df_late <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg+25)-3, (peak_avg+25)+3))),
         season = "late") %>% 
  unnest(cols = c(data))

plot_df <- bind_rows(list(plot_df_early, plot_df_peak, plot_df_late)) %>% 
  mutate(date_start = case_when(
    season == "peak" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 3, "%d %b"),
    season == "early" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 28, "%d %b"),
    season == "late" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 22, "%d %b")
  )) %>% 
  mutate(date_end = case_when(
    season == "peak" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 3, "%d %b"),
    season == "early" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 22, "%d %b"),
    season == "late" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 28, "%d %b")
  )) %>% 
  mutate(season = forcats::fct_relevel(season, "early", "peak", "late"),
         season = forcats::fct_recode(season, "peak haul out - 25 days" = "early",
                             "peak haul out" = "peak",
                             "peak haul out + 25 days" = "late"))

p_temp <- ggplot(plot_df, 
                 aes(temp2*27, ho_prob, color = age_sex, fill = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.1, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, size = 0.3, alpha = 0.1,
              show.legend = FALSE) +
  scale_fill_d3() +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1),
                                      direction = "horizontal")) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  facet_wrap(. ~ season + date_start + date_end, scales = "free_x",
             labeller = label_glue('{season}\n{date_start} - {date_end}')) +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df, aes(wind*10, ho_prob, color = age_sex, fill = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.1, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, size = 0.3, alpha = 0.1,
              show.legend = FALSE) +
  scale_fill_d3() +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 0.75),
                                      direction = "horizontal")) +
  facet_wrap(. ~ season + date_start + date_end, scales = "free_x",
             labeller = label_glue('{season}\n{date_start} - {date_end}')) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  xlab("total wind (m/s)") + ylab("haul-out probability")

@@

p_temp / p_wind +
  plot_annotation(
    title = 'The influence of temperature and wind on predicted ribbon seal \nhaul-out behavior varies within season',
    subtitle = 'data are subset to only include local solar hours between 0800 and 1600',
    caption = 'Solid lines represent linear fits to the predicted values and are \nprovided to better indicate marginal patterns.')
```

@@

```{r spottedHOwx,fig.height = 5.5, include=TRUE, fig.id = "spottedHOwx", fig.cap="Variability in the predicted haul-out probability of spotted seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

plot_df <- spotted_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label = TRUE),
         day = lubridate::day(date)) %>% 
  filter(age_sex != "YOUNG OF YEAR",
         between(solar_hour, 8,16)) %>% 
  group_by(age_sex)

plot_df_early <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg-25)-3, (peak_avg-25)+3))),
         season = "early") %>% 
  unnest(cols = c(data))

plot_df_peak <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg)-3, (peak_avg)+3))),
         season = "peak") %>% 
  unnest(cols = c(data))

plot_df_late <- plot_df %>% tidyr::nest() %>% 
  rowwise() %>% 
  mutate(peak_idx = which.max(data$ho_prob),
         peak_day = data$yday[peak_idx]) %>% ungroup() %>% 
  mutate(peak_avg = as.integer(mean(peak_day))) %>% 
  rowwise() %>% 
  mutate(data = list(data %>% filter(between(yday,(peak_avg+25)-3, (peak_avg+25)+3))),
         season = "late") %>% 
  unnest(cols = c(data))

plot_df <- bind_rows(list(plot_df_early, plot_df_peak, plot_df_late)) %>% 
  mutate(date_start = case_when(
    season == "peak" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 3, "%d %b"),
    season == "early" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 28, "%d %b"),
    season == "late" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 22, "%d %b")
  )) %>% 
  mutate(date_end = case_when(
    season == "peak" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 3, "%d %b"),
    season == "early" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") - 22, "%d %b"),
    season == "late" ~ format(lubridate::as_date(peak_avg, origin = "2015-01-01") + 28, "%d %b")
  )) %>% 
  mutate(season = forcats::fct_relevel(season, "early", "peak", "late"),
         season = forcats::fct_recode(season, "peak haul out - 25 days" = "early",
                             "peak haul out" = "peak",
                             "peak haul out + 25 days" = "late"))

p_temp <- ggplot(plot_df, 
                 aes(temp2*27, ho_prob, color = age_sex, fill = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.1, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, size = 0.3, alpha = 0.1,
              show.legend = FALSE) +
  scale_fill_d3() +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1),
                                      direction = "horizontal")) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  facet_wrap(. ~ season + date_start + date_end, scales = "free_x",
             labeller = label_glue('{season}\n{date_start} - {date_end}')) +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df, aes(wind*10, ho_prob, color = age_sex, fill = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.1, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, size = 0.3, alpha = 0.1,
              show.legend = FALSE) +
  scale_fill_d3() +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 0.75),
                                      direction = "horizontal")) +
  facet_wrap(. ~ season + date_start + date_end, scales = "free_x",
             labeller = label_glue('{season}\n{date_start} - {date_end}')) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  xlab("total wind (m/s)") + ylab("haul-out probability")

@@

p_temp / p_wind +
  plot_annotation(
    title = 'The influence of temperature and wind on predicted spotted seal \nhaul-out behavior varies within season',
    subtitle = 'data are subset to only include local solar hours between 0800 and 1600',
    caption = 'Solid lines represent linear fits to the predicted values and are \nprovided to better indicate marginal patterns.')
```
